% !TEX root = proposal.tex

\acknowledgements{

\vspace{-0.3in}

I am eternally grateful to my parents, Andrew and Margaret Adrian, for their
unwavering support and the privilege to be in a position where hard work can
lead to success, as well as to my brother, Alex Adrian. I'd like to thank my
advisor, Professor J.~Alex Halderman. I've learned tremendous amount over the
last five years, and I will never forget my time as your student. I'd also
like to thank Peter Honeyman and the rest of my commitee---Chris Peikert and
Florian Schaub---for pushing me towards the door. I am especially thankful to
Zakir Durumeric, for all projects we worked on together, and to Pat
Pannuto---had we not lived together, I likely would have left the program. I
am similarly grateful to Bradford Campbell, Noah Klugman, and Kyle Lady for
keeping me sane. I appreciate the support and advice I received from
Professors Prabal Dutta, and Brian Noble, as well as the comradery and late
nights with labmates and coauthors: Nimrod Aviram, Matt Bernhard, Branden
Ghena, Alex Holland, Will Huang, James Kasten, Deepak Kumar, Allison
McDonald, Ariana Mirian, Deepika Natarajan, Drew Springall, Ben VanderSloot,
Eric Wustrow, and many others. I credit Professor Robert Dick, and his
students David Bild and Yue Liu, for opening up the world of research to me
in undergrad. I am thankful to many known friends and trusted agents:
Charlotte Campbell, Elson Liu, Lane Powell, and Ethan Stark; Adrienne Porter
Felt, Emily Stark, and Ryan Sleevi; Adam Goodman, Jon Oberheide, and Dug
Song; Zach Lanier, Thomas Ptacek, and George Tankersley; Bennett Black,
Veronica Long, Kyle Schiller and Tristan Weber; Michael Moore; and Dave
Corcoran, Chris Dzombak, Brian Kelly, Andrew Sardone, and everyone at Censys.

The work in this dissertation was supported in part by the National Science
Foundation and the Alfred P. Sloan Foundation, with additional support from
the Mozilla Foundation, Supermicro, Google, Cisco, and the Morris Wellman
Professorship.

Forever Go Blue!

}
