\documentclass[doublespace]{styles/rackham-thesis}

\usepackage[linktocpage=true]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb, amsthm, xfrac}
\usepackage{booktabs}
\usepackage{lipsum}
\usepackage{flafter}

\include{includes/thesis-headers}

\begin{document}

\title{Using Large-Scale Empirical Methods to Understand Fragile Cryptographic Ecosystems}
\author{David Adrian}
\degree{Doctor of Philosophy}
\department{Computer Science and Engineering}
\committee{
  Professor J.~Alex Halderman, Chair\\
  Research Professor Peter Honeyman\\
  Associate Professor Chris Peikert\\
  Assistant Professor Florian Schaub\\
}

\maketitle

\input{dedication}
\input{acknowledgements}

\tableofcontents
\listoffigures
\listoftables
\listofappendices

\input{abstract}

\chapter{Introduction}
\input{proposal/intro}

\chapter{Improving Measurement}
\label{chapter:zippier}
\input{papers/zippy/paper/paper}
\chapter{Measuring Diffie-Hellman}
\label{chapter:subgroup}
\input{papers/subgroup/paper/subgroup}
\chapter{Measuring Export-Grade Key Exchange}
\label{chapter:logjam}
\input{papers/logjam/paper/logjam}
\chapter{Measuring Export-Grade Symmetric Cryptography}
\label{chapter:drown}
\input{papers/drown/paper/drown}

\chapter{Conclusion and Future Work}
\label{chapter:conclusion}

\textit{This is an outline}

List of how the empirical methods discussed earlier had impact
(pull-out relevant results).

Did any of this research go anywhere beyond that? Yes, it had a clear impact
on TLS 1.3.

\section{TLS 1.3}

Point out that the ``informing future protocol design'' is effectively
``making TLS 1.3 better''. This is the ``improve the security of the Internet
in the future'' part.

Call out explicit TLS 1.3 design decisions based on results described in this
dissertation. Perhaps some of this should be interleaved into the earlier
sections?

\section{Engineering Challenges}

For not being a ``systems'' field, there is still an absurd amount of
engineering that largely goes unacknowledged, in order to write good
measurement papers. Discuss some of these examples.

Is this a fundamental state of being of the methodology, or are we doing something wrong?

\section{Broader applicability of empirical methods}

Security driven by data. Is any of this relevant to users who are just trying
to secure this own networks?

Extending measurement from aggregations and ecosystems to behavior of
individual hosts at global scale. Can we track individual hosts appearing and
disappearing, and map changes to configuration in real-time?

Should empiricism be part of more traditional cryptography education and research?

Some word of warning about how measurement doesn't solve all our problems,
and how measuring the wrong things makes things worse, \eg Robert McNamera's
use of body count as a metric during the Vietnam War.

% What the fuck does anything have to do with Vietnam?

\bibliographystyle{plain}
\bibliography{refs}

\end{document}
