\documentclass{thesis-umich}
\usepackage[margin=1in]{geometry}
\usepackage{mathptmx}\renewcommand{\ttdefault}{cmtt}
\usepackage{graphicx}

%\usepackage{microtype}
\usepackage{xspace}
\usepackage[compact]{titlesec}
\usepackage{color}
\definecolor{azure}{rgb}{0.16, 0.32, 0.75}
%\usepackage[%
%    breaklinks=true,colorlinks=true,urlcolor=azure,
%    citecolor=black,pdftex]{hyperref}
%\thispagestyle{empty}\pagestyle{plain}
\usepackage{url}\urlstyle{rm}

\usepackage{amsmath,amssymb,amsbsy}

\usepackage[margin=0pt]{caption}
\usepackage{algorithm, algorithmic}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\usepackage{color}
\usepackage{verbatim}
%\usepackage[margin=1in]{geometry}
%\usepackage{watermark}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{cite}
\usepackage{wasysym}
\usepackage{paralist}
\usepackage{balance}
\usepackage{tabularx}
\usepackage[protrusion=true,expansion=true,kerning]{microtype}
\usepackage{multirow}
\usepackage{placeins}
\usepackage{pgfplots}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{adjustbox}
\usepackage{bigstrut}
%\usepackage{subcaption}
\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\newcommand{\twolinecell}[2][r]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\newcommand{\twollinecell}[2][r]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}


\newcommand*{\refname}{Bibliography}

\usepackage{float}

\newcommand{\ind}{\hspace*{1em}}
\newif\ifex\extrue % for extended version
\newif\ifredact\redactfalse
\newcommand{\re}[1]{{#1}}
\newcommand{\subpar}[1]{\medskip\noindent\textsl{#1}\enspace}
\newcommand{\update}[1]{{\color{black}#1}\xspace}
\newcommand{\mycaption}[2]{\caption{\textbf{#1.} #2}}
\newcommand{\hostfont}[1]{\texttt{#1}}
\newcommand{\starttls}{STARTTLS\@\xspace}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\color{red}{\ding{53}}}

\usepackage{compactbib}
%\usepackage[compress]{cite}
%\renewcommand{\bibliorefname}[1]{\subsection*{#1}}
\renewcommand{\paragraph}[1]{\medskip\noindent\textbf{#1}\,\,\,}

\newcommand{\todo}[1]{{\color{red}{\textbf{\em [TODO: #1]}}}\xspace}
\newcommand{\TODO}[1]{\todo{#1}}

\begin{document}

\chapter{Introduction}

General introduction. Broad claims. Thesis statement.

No one will actually read past here.

\section{First Paper Class}

Introduction and claim group. Cite something~\cite{weakkeys12}.

\subsection{Paper One}

Now we're cooking with gas!

\paragraph{Important Note}
This paper is more exciting and gets some breakout points.

\paragraph{Another Note}
All this stuff is made up anyway.

\subsection{Another Paper}

We talked about some stuff, and now we're gonna fly at left-field.

\section{Second Paper Group}

Let's cite a bunch of things that are related, then talk about them!

\subsection{A paper}

I published a thing.

\subsection{Analyzing the HTTPS Certificate Ecosystem}

And another thing

\paragraph{Note}
Another note.

\paragraph{Site Certificates}
A second note.

\chapter{Paper 1}
\input{zmap/zpaper}
\chapter{Paper 2}
\input{censys/zsearch}
\chapter{Paper 3}
\input{psandqs/paper3}
\chapter{Paper 4}
\input{httpsecosystem/imc13}
\chapter{Paper 5}
\input{mail/paper}
\chapter{Paper 6}
\input{heartbleed/heartbleed}

\chapter{Conclusion and Future Work}

Concluding thoughts.

Future work

\section{Future Idea A}
Graduate.

\section{Future Idea B}
Startup.

\section{Shaping security behavior}
Cash money.

\bibliographystyle{plain}
\bibliography{refs}

\end{document}